---
title: "FA7_Dizon_STATS"
author: "Kiana Marie Dizon"
output: html_document
date: "2024-09-27"
---

## Introduction

#### This analysis aims to compare the means of two groups (those using the invisibility cloak and those not using it) in terms of their "Mischief" scores.


```{r 1}
library(dplyr)
library(ggplot2)

# Create the dataset
data <- data.frame(
  Participant = 1:24,
  Cloak = c(rep(0, 12), rep(1, 12)),  # 0 = no cloak, 1 = cloak
  Mischief = c(3, 1, 5, 4, 6, 4, 6, 2, 0, 5, 4, 5,
               4, 3, 6, 6, 8, 5, 5, 4, 2, 5, 7, 5)
)

# Display the data
head(data)
```
## Assumptions of the Independent Samples t-test
#### Assumption 1: Independence of Observations
#### The participants in each group are independent. There should be no relationship between the individuals in the different groups.

#### Assumption 2: Normality
#### The distribution of scores in each group should be approximately normally distributed. This can be checked using Shapiro-Wilk test.

```{r 2}
shapiro_test_no_cloak <- shapiro.test(data$Mischief[data$Cloak == 0])
shapiro_test_with_cloak <- shapiro.test(data$Mischief[data$Cloak == 1])

shapiro_test_no_cloak
shapiro_test_with_cloak
```

#### Assumption 3: Homogeneity of Variance
#### The variances in each group should be approximately equal. This can be checked using Levene's Test.


```{r 3}
# Install and load car package for Levene's Test
if (!requireNamespace("car", quietly = TRUE)) {
    install.packages("car")
}
library(car)

# Levene's Test
levene_test <- leveneTest(Mischief ~ as.factor(Cloak), data = data)
levene_test
```
#### Assumption 4: Scale of Measurement
#### The dependent variable (Mischief scores) is measured on a continuous scale.

#### Assumption 5: Random Sampling
#### The sample should be randomly selected from the population.

```{r 4}
# Independent samples t-test
t_test_result <- t.test(Mischief ~ Cloak, data = data, var.equal = TRUE)  # Change to var.equal = FALSE if variances are unequal
t_test_result
```


```{r 5}
# Extract results
t_value <- t_test_result$statistic
p_value <- t_test_result$p.value
mean_no_cloak <- mean(data$Mischief[data$Cloak == 0])
mean_with_cloak <- mean(data$Mischief[data$Cloak == 1])
confidence_interval <- t_test_result$conf.int

# Summarize results
cat("T-Value:", t_value, "\n")
cat("P-Value:", p_value, "\n")
cat("Mean Mischief (No Cloak):", mean_no_cloak, "\n")
cat("Mean Mischief (With Cloak):", mean_with_cloak, "\n")
cat("95% Confidence Interval:", confidence_interval, "\n")
```

## Conclusion
#### Based on the independent samples t-test, we observed a t-value of r round(t_value, -1.71) with a p-value of r round(p_value, 0.10). The mean mischief score for participants without the cloak was r round(mean_no_cloak, 3.75), while for those with the cloak, it was r round(mean_with_cloak, 5). The 95% confidence interval for the difference in means is r round(confidence_interval[1], -2.76) to r round(confidence_interval[2], 0.26).

#### If the p-value is less than the significance level (typically 0.05), we reject the null hypothesis, concluding that there is a statistically significant difference in the mischief scores between the two groups.
